---
title: "Simulation Summary"
output: html_document
---

```{r, include = FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
### Set working directory to correct location
cdir <- stringr::str_split(getwd(), "/")[[1]]
udir <- cdir[1:which(cdir == "State-Space-Methods")]
setwd(paste(udir, collapse = "/"))

fct_case_when <- function(...) {
  args <- as.list(match.call())
  levels <- sapply(args[-1], function(f) f[[3]])  # extract RHS of formula
  levels <- levels[!is.na(levels)]
  factor(dplyr::case_when(...), levels=levels)
}
knitr::opts_chunk$set(echo = FALSE)
```


```{r}
flist <- list.files("Simulations/SimOutReform")
flist <- flist[grepl("Itout", flist)]
flist <- flist[c(3, 1, 2, 4, 6, 5, 7, 8)]
B <- c(4, 2, -1)
```




```{r, results='asis'}
find <- 1
Itout <- readRDS(paste("Simulations/SimOutReform/",flist[find], sep = "")) 
MedEst <- apply(Itout$Estimate, 1:2, median, na.rm = TRUE)
MedBias <- MedEst - B
VarEst <- apply(Itout$Estimate, 1:2, var, na.rm = TRUE)
Coverage <- apply(Itout$UCL >= B & Itout$LCL <= B, 1:2, mean, na.rm = TRUE) 
colnames(MedBias) <- colnames(VarEst) <- colnames(Coverage) <- c("LME", "LME AR(1)", "SS Orig", "SS Bayes", "SS Part 2", "SS Part 4", "SS Part 10") 

vvec <- flist[find] %>% 
  gsub("Itout ", "", .) %>%
  gsub(".RDS", "", .) %>%
  str_split(" ") %>%
  .[[1]] %>%
  gsub("_", ".", .)


```



## $\sigma^2_\varepsilon =$ `r vvec[1]`, $\sigma^2_\eta =$ `r vvec[2]`

### Coverage

```{r}
Coverage %>%
  kable() %>% kable_styling()
```


### Bias Plot

```{r, fig.height=7}
Bias <- Itout$Estimate - B
Var1Bias <- t(Bias[1,,])
Var2Bias <- t(Bias[2,,])
Var3Bias <- t(Bias[3,,])
VarBias <- data.frame(rbind(Var1Bias, Var2Bias, Var3Bias))

VarBias$Variable <- sort(rep(paste("B", 1:3, sep = ""), dim(Bias)[3]))



colnames(VarBias) <-  c("LME", "LME AR(1)", "SS Orig", "SS Bayes", "SS Part 2", "SS Part 4", "SS Part 10",  "Variable") 

VarBias %>%
  data.frame() %>%
  gather("Method", "Bias", `LME`:`SS.Part.10`) %>%
  mutate(Method = fct_case_when(
    Method == "LME" ~ "LME",
    Method == "LME.AR.1." ~ "LME AR(1)",
    Method == "SS.Orig" ~ "SS Original",
    Method == "SS.Bayes" ~ "SS Bayes",
    Method == "SS.Part.2" ~ "SS Partition k = 2",
    Method == "SS.Part.4" ~ "SS Partition k = 4",
    Method == "SS.Part.10" ~ "SS Partition k = 10"
  )) %>%
  filter(Variable == "B1") %>%
  group_by(Method) %>%
  filter(Bias < quantile(Bias, .99) & Bias > quantile(Bias, .01)) %>%
  ungroup(Method) %>%
  ggplot(aes(x = Method, y = Bias, fill = Method)) + 
  geom_violin() + 
  stat_summary(fun=median, geom="point", shape=18, size=5) +
  facet_grid(Variable ~ .) +
  theme(axis.text.x = element_text(angle = -45, vjust = .99, hjust = 0), legend.position = "none")
```

### CI Length Plot

```{r, fig.height=7}
CIlen <- Itout$UCL -Itout$LCL
CIlen1 <- t(CIlen[1,,])
CIlen2 <- t(CIlen[2,,])
CIlen3 <- t(CIlen[3,,])
CIlenTot <- data.frame(rbind(CIlen1, CIlen2, CIlen3))

CIlenTot$Variable <- sort(rep(paste("B", 1:3, sep = ""), dim(Bias)[3]))



colnames(CIlenTot) <-  c("LME", "LME AR(1)", "SS Orig", "SS Bayes", "SS Part 2", "SS Part 4", "SS Part 10",  "Variable") 

CIlenTot %>%
  data.frame() %>%
  gather("Method", "CI.Length", `LME`:`SS.Part.10`) %>%
  mutate(Method = fct_case_when(
    Method == "LME" ~ "LME",
    Method == "LME.AR.1." ~ "LME AR(1)",
    Method == "SS.Orig" ~ "SS Original",
    Method == "SS.Bayes" ~ "SS Bayes",
    Method == "SS.Part.2" ~ "SS Partition k = 2",
    Method == "SS.Part.4" ~ "SS Partition k = 4",
    Method == "SS.Part.10" ~ "SS Partition k = 10"
  )) %>%
  filter(Variable == "B1")%>%
  filter(!is.na(CI.Length)) %>%
      group_by(Method) %>%
  filter(CI.Length < quantile(CI.Length, .99) & CI.Length > quantile(CI.Length, .01)) %>%
  ungroup(Method) %>%
  ggplot(aes(x = Method, y = CI.Length, fill = Method)) + 
  geom_violin() + 
  stat_summary(fun=median, geom="point", shape=18, size=5) +
  facet_grid(Variable ~ .) +
  theme(axis.text.x = element_text(angle = -45, vjust = .99, hjust = 0), legend.position = "none")
```


















```{r, results='asis'}
find <- 2
Itout <- readRDS(paste("Simulations/SimOutReform/",flist[find], sep = "")) 
MedEst <- apply(Itout$Estimate, 1:2, median, na.rm = TRUE)
MedBias <- MedEst - B
VarEst <- apply(Itout$Estimate, 1:2, var, na.rm = TRUE)
Coverage <- apply(Itout$UCL >= B & Itout$LCL <= B, 1:2, mean, na.rm = TRUE) 
colnames(MedBias) <- colnames(VarEst) <- colnames(Coverage) <- c("LME", "LME AR(1)", "SS Orig", "SS Bayes", "SS Part 2", "SS Part 4", "SS Part 10") 

vvec <- flist[find] %>% 
  gsub("Itout ", "", .) %>%
  gsub(".RDS", "", .) %>%
  str_split(" ") %>%
  .[[1]] %>%
  gsub("_", ".", .)


```


## $\sigma^2_\varepsilon =$ `r vvec[1]`, $\sigma^2_\eta =$ `r vvec[2]`

### Coverage

```{r}
Coverage %>%
  kable() %>% kable_styling()
```


### Bias Plot

```{r, fig.height=7}
Bias <- Itout$Estimate - B
Var1Bias <- t(Bias[1,,])
Var2Bias <- t(Bias[2,,])
Var3Bias <- t(Bias[3,,])
VarBias <- data.frame(rbind(Var1Bias, Var2Bias, Var3Bias))

VarBias$Variable <- sort(rep(paste("B", 1:3, sep = ""), dim(Bias)[3]))



colnames(VarBias) <-  c("LME", "LME AR(1)", "SS Orig", "SS Bayes", "SS Part 2", "SS Part 4", "SS Part 10",  "Variable") 

VarBias %>%
  data.frame() %>%
  gather("Method", "Bias", `LME`:`SS.Part.10`) %>%
  mutate(Method = fct_case_when(
    Method == "LME" ~ "LME",
    Method == "LME.AR.1." ~ "LME AR(1)",
    Method == "SS.Orig" ~ "SS Original",
    Method == "SS.Bayes" ~ "SS Bayes",
    Method == "SS.Part.2" ~ "SS Partition k = 2",
    Method == "SS.Part.4" ~ "SS Partition k = 4",
    Method == "SS.Part.10" ~ "SS Partition k = 10"
  )) %>%
  filter(Variable == "B1")%>%
    group_by(Method) %>%
  filter(Bias < quantile(Bias, .99) & Bias > quantile(Bias, .01)) %>%
  ungroup(Method) %>%
  ggplot(aes(x = Method, y = Bias, fill = Method)) + 
  geom_violin() + 
  stat_summary(fun=median, geom="point", shape=18, size=5) +
  facet_grid(Variable ~ .) +
  theme(axis.text.x = element_text(angle = -45, vjust = .99, hjust = 0), legend.position = "none")
```


### CI Length Plot

```{r, fig.height=7}
CIlen <- Itout$UCL -Itout$LCL
CIlen1 <- t(CIlen[1,,])
CIlen2 <- t(CIlen[2,,])
CIlen3 <- t(CIlen[3,,])
CIlenTot <- data.frame(rbind(CIlen1, CIlen2, CIlen3))

CIlenTot$Variable <- sort(rep(paste("B", 1:3, sep = ""), dim(Bias)[3]))



colnames(CIlenTot) <-  c("LME", "LME AR(1)", "SS Orig", "SS Bayes", "SS Part 2", "SS Part 4", "SS Part 10",  "Variable") 

CIlenTot %>%
  data.frame() %>%
  gather("Method", "CI.Length", `LME`:`SS.Part.10`) %>%
  mutate(Method = fct_case_when(
    Method == "LME" ~ "LME",
    Method == "LME.AR.1." ~ "LME AR(1)",
    Method == "SS.Orig" ~ "SS Original",
    Method == "SS.Bayes" ~ "SS Bayes",
    Method == "SS.Part.2" ~ "SS Partition k = 2",
    Method == "SS.Part.4" ~ "SS Partition k = 4",
    Method == "SS.Part.10" ~ "SS Partition k = 10"
  ))%>%
  filter(Variable == "B1") %>%
  filter(!is.na(CI.Length)) %>%
      group_by(Method) %>%
  filter(CI.Length < quantile(CI.Length, .99) & CI.Length > quantile(CI.Length, .01)) %>%
  ungroup(Method) %>%
  ggplot(aes(x = Method, y = CI.Length, fill = Method)) + 
  geom_violin() + 
  stat_summary(fun=median, geom="point", shape=18, size=5) +
  facet_grid(Variable ~ .) +
  theme(axis.text.x = element_text(angle = -45, vjust = .99, hjust = 0), legend.position = "none")
```












```{r, results='asis'}
find <- 3
Itout <- readRDS(paste("Simulations/SimOutReform/",flist[find], sep = "")) 
MedEst <- apply(Itout$Estimate, 1:2, median, na.rm = TRUE)
MedBias <- MedEst - B
VarEst <- apply(Itout$Estimate, 1:2, var, na.rm = TRUE)
Coverage <- apply(Itout$UCL >= B & Itout$LCL <= B, 1:2, mean, na.rm = TRUE) 
colnames(MedBias) <- colnames(VarEst) <- colnames(Coverage) <- c("LME", "LME AR(1)", "SS Orig", "SS Bayes", "SS Part 2", "SS Part 4", "SS Part 10") 

vvec <- flist[find] %>% 
  gsub("Itout ", "", .) %>%
  gsub(".RDS", "", .) %>%
  str_split(" ") %>%
  .[[1]] %>%
  gsub("_", ".", .)


```


## $\sigma^2_\varepsilon =$ `r vvec[1]`, $\sigma^2_\eta =$ `r vvec[2]`

### Coverage

```{r}
Coverage %>%
  kable() %>% kable_styling()
```


### Bias Plot

```{r, fig.height=7}
Bias <- Itout$Estimate - B
Var1Bias <- t(Bias[1,,])
Var2Bias <- t(Bias[2,,])
Var3Bias <- t(Bias[3,,])
VarBias <- data.frame(rbind(Var1Bias, Var2Bias, Var3Bias))

VarBias$Variable <- sort(rep(paste("B", 1:3, sep = ""), dim(Bias)[3]))



colnames(VarBias) <-  c("LME", "LME AR(1)", "SS Orig", "SS Bayes", "SS Part 2", "SS Part 4", "SS Part 10",  "Variable") 

VarBias %>%
  data.frame() %>%
  gather("Method", "Bias", `LME`:`SS.Part.10`) %>%
  mutate(Method = fct_case_when(
    Method == "LME" ~ "LME",
    Method == "LME.AR.1." ~ "LME AR(1)",
    Method == "SS.Orig" ~ "SS Original",
    Method == "SS.Bayes" ~ "SS Bayes",
    Method == "SS.Part.2" ~ "SS Partition k = 2",
    Method == "SS.Part.4" ~ "SS Partition k = 4",
    Method == "SS.Part.10" ~ "SS Partition k = 10"
  )) %>%
  filter(Variable == "B1")%>%
    group_by(Method) %>%
  filter(Bias < quantile(Bias, .99) & Bias > quantile(Bias, .01)) %>%
  ungroup(Method) %>%
  ggplot(aes(x = Method, y = Bias, fill = Method)) + 
  geom_violin() + 
  stat_summary(fun=median, geom="point", shape=18, size=5) +
  facet_grid(Variable ~ .) +
  theme(axis.text.x = element_text(angle = -45, vjust = .99, hjust = 0), legend.position = "none")
```



### CI Length Plot

```{r, fig.height=7}
CIlen <- Itout$UCL -Itout$LCL
CIlen1 <- t(CIlen[1,,])
CIlen2 <- t(CIlen[2,,])
CIlen3 <- t(CIlen[3,,])
CIlenTot <- data.frame(rbind(CIlen1, CIlen2, CIlen3))

CIlenTot$Variable <- sort(rep(paste("B", 1:3, sep = ""), dim(Bias)[3]))



colnames(CIlenTot) <-  c("LME", "LME AR(1)", "SS Orig", "SS Bayes", "SS Part 2", "SS Part 4", "SS Part 10",  "Variable") 

CIlenTot %>%
  data.frame() %>%
  gather("Method", "CI.Length", `LME`:`SS.Part.10`) %>%
  mutate(Method = fct_case_when(
    Method == "LME" ~ "LME",
    Method == "LME.AR.1." ~ "LME AR(1)",
    Method == "SS.Orig" ~ "SS Original",
    Method == "SS.Bayes" ~ "SS Bayes",
    Method == "SS.Part.2" ~ "SS Partition k = 2",
    Method == "SS.Part.4" ~ "SS Partition k = 4",
    Method == "SS.Part.10" ~ "SS Partition k = 10"
  )) %>%
  filter(Variable == "B1")%>%
  filter(!is.na(CI.Length)) %>%
      group_by(Method) %>%
  filter(CI.Length < quantile(CI.Length, .99) & CI.Length > quantile(CI.Length, .01)) %>%
  ungroup(Method) %>%
  ggplot(aes(x = Method, y = CI.Length, fill = Method)) + 
  geom_violin() + 
  stat_summary(fun=median, geom="point", shape=18, size=5) +
  facet_grid(Variable ~ .) +
  theme(axis.text.x = element_text(angle = -45, vjust = .99, hjust = 0), legend.position = "none")
```










```{r, results='asis'}
find <- 4
Itout <- readRDS(paste("Simulations/SimOutReform/",flist[find], sep = "")) 
MedEst <- apply(Itout$Estimate, 1:2, median, na.rm = TRUE)
MedBias <- MedEst - B
VarEst <- apply(Itout$Estimate, 1:2, var, na.rm = TRUE)
Coverage <- apply(Itout$UCL >= B & Itout$LCL <= B, 1:2, mean, na.rm = TRUE) 
colnames(MedBias) <- colnames(VarEst) <- colnames(Coverage) <- c("LME", "LME AR(1)", "SS Orig", "SS Bayes", "SS Part 2", "SS Part 4", "SS Part 10") 

vvec <- flist[find] %>% 
  gsub("Itout ", "", .) %>%
  gsub(".RDS", "", .) %>%
  str_split(" ") %>%
  .[[1]] %>%
  gsub("_", ".", .)


```

## $\sigma^2_\varepsilon =$ `r vvec[1]`, $\sigma^2_\eta =$ `r vvec[2]`

### Coverage

```{r}
Coverage %>%
  kable() %>% kable_styling()
```


### Bias Plot

```{r, fig.height=7}
Bias <- Itout$Estimate - B
Var1Bias <- t(Bias[1,,])
Var2Bias <- t(Bias[2,,])
Var3Bias <- t(Bias[3,,])
VarBias <- data.frame(rbind(Var1Bias, Var2Bias, Var3Bias))

VarBias$Variable <- sort(rep(paste("B", 1:3, sep = ""), dim(Bias)[3]))



colnames(VarBias) <-  c("LME", "LME AR(1)", "SS Orig", "SS Bayes", "SS Part 2", "SS Part 4", "SS Part 10",  "Variable") 

VarBias %>%
  data.frame() %>%
  gather("Method", "Bias", `LME`:`SS.Part.10`) %>%
  mutate(Method = fct_case_when(
    Method == "LME" ~ "LME",
    Method == "LME.AR.1." ~ "LME AR(1)",
    Method == "SS.Orig" ~ "SS Original",
    Method == "SS.Bayes" ~ "SS Bayes",
    Method == "SS.Part.2" ~ "SS Partition k = 2",
    Method == "SS.Part.4" ~ "SS Partition k = 4",
    Method == "SS.Part.10" ~ "SS Partition k = 10"
  )) %>%
  filter(Variable == "B1")%>%
    group_by(Method) %>%
  filter(Bias < quantile(Bias, .99) & Bias > quantile(Bias, .01)) %>%
  ungroup(Method) %>%
  ggplot(aes(x = Method, y = Bias, fill = Method)) + 
  geom_violin() + 
  stat_summary(fun=median, geom="point", shape=18, size=5) +
  facet_grid(Variable ~ .) +
  theme(axis.text.x = element_text(angle = -45, vjust = .99, hjust = 0), legend.position = "none")
```

### CI Length Plot

```{r, fig.height=7}
CIlen <- Itout$UCL -Itout$LCL
CIlen1 <- t(CIlen[1,,])
CIlen2 <- t(CIlen[2,,])
CIlen3 <- t(CIlen[3,,])
CIlenTot <- data.frame(rbind(CIlen1, CIlen2, CIlen3))

CIlenTot$Variable <- sort(rep(paste("B", 1:3, sep = ""), dim(Bias)[3]))



colnames(CIlenTot) <-  c("LME", "LME AR(1)", "SS Orig", "SS Bayes", "SS Part 2", "SS Part 4", "SS Part 10",  "Variable") 

CIlenTot %>%
  data.frame() %>%
  gather("Method", "CI.Length", `LME`:`SS.Part.10`) %>%
  mutate(Method = fct_case_when(
    Method == "LME" ~ "LME",
    Method == "LME.AR.1." ~ "LME AR(1)",
    Method == "SS.Orig" ~ "SS Original",
    Method == "SS.Bayes" ~ "SS Bayes",
    Method == "SS.Part.2" ~ "SS Partition k = 2",
    Method == "SS.Part.4" ~ "SS Partition k = 4",
    Method == "SS.Part.10" ~ "SS Partition k = 10"
  )) %>%
  filter(Variable == "B1")%>%
      group_by(Method)%>%
  filter(!is.na(CI.Length))  %>%
  filter(CI.Length < quantile(CI.Length, .99) & CI.Length > quantile(CI.Length, .01)) %>%
  ungroup(Method) %>%
  ggplot(aes(x = Method, y = CI.Length, fill = Method)) + 
  geom_violin() + 
  stat_summary(fun=median, geom="point", shape=18, size=5) +
  facet_grid(Variable ~ .) +
  theme(axis.text.x = element_text(angle = -45, vjust = .99, hjust = 0), legend.position = "none")
```














```{r, results='asis'}
find <- 5
Itout <- readRDS(paste("Simulations/SimOutReform/",flist[find], sep = "")) 
MedEst <- apply(Itout$Estimate, 1:2, median, na.rm = TRUE)
MedBias <- MedEst - B
VarEst <- apply(Itout$Estimate, 1:2, var, na.rm = TRUE)
Coverage <- apply(Itout$UCL >= B & Itout$LCL <= B, 1:2, mean, na.rm = TRUE) 
colnames(MedBias) <- colnames(VarEst) <- colnames(Coverage) <- c("LME", "LME AR(1)", "SS Orig", "SS Bayes", "SS Part 2", "SS Part 4", "SS Part 10") 

vvec <- flist[find] %>% 
  gsub("Itout ", "", .) %>%
  gsub(".RDS", "", .) %>%
  str_split(" ") %>%
  .[[1]] %>%
  gsub("_", ".", .)


```

## $\sigma^2_\varepsilon =$ `r vvec[1]`, $\sigma^2_\eta =$ `r vvec[2]`


### Coverage

```{r}
Coverage %>%
  kable() %>% kable_styling()
```


### Bias Plot

```{r, fig.height=7}
Bias <- Itout$Estimate - B
Var1Bias <- t(Bias[1,,])
Var2Bias <- t(Bias[2,,])
Var3Bias <- t(Bias[3,,])
VarBias <- data.frame(rbind(Var1Bias, Var2Bias, Var3Bias))

VarBias$Variable <- sort(rep(paste("B", 1:3, sep = ""), dim(Bias)[3]))



colnames(VarBias) <-  c("LME", "LME AR(1)", "SS Orig", "SS Bayes", "SS Part 2", "SS Part 4", "SS Part 10",  "Variable") 

VarBias %>%
  data.frame() %>%
  gather("Method", "Bias", `LME`:`SS.Part.10`) %>%
  mutate(Method = fct_case_when(
    Method == "LME" ~ "LME",
    Method == "LME.AR.1." ~ "LME AR(1)",
    Method == "SS.Orig" ~ "SS Original",
    Method == "SS.Bayes" ~ "SS Bayes",
    Method == "SS.Part.2" ~ "SS Partition k = 2",
    Method == "SS.Part.4" ~ "SS Partition k = 4",
    Method == "SS.Part.10" ~ "SS Partition k = 10"
  )) %>%
  filter(Variable == "B1")%>%
    group_by(Method) %>%
  filter(Bias < quantile(Bias, .99) & Bias > quantile(Bias, .01)) %>%
  ungroup(Method) %>%
  ggplot(aes(x = Method, y = Bias, fill = Method)) + 
  geom_violin() + 
  stat_summary(fun=median, geom="point", shape=18, size=5) +
  facet_grid(Variable ~ .) +
  theme(axis.text.x = element_text(angle = -45, vjust = .99, hjust = 0), legend.position = "none")
```

### CI Length Plot

```{r, fig.height=7}
CIlen <- Itout$UCL -Itout$LCL
CIlen1 <- t(CIlen[1,,])
CIlen2 <- t(CIlen[2,,])
CIlen3 <- t(CIlen[3,,])
CIlenTot <- data.frame(rbind(CIlen1, CIlen2, CIlen3))

CIlenTot$Variable <- sort(rep(paste("B", 1:3, sep = ""), dim(Bias)[3]))



colnames(CIlenTot) <-  c("LME", "LME AR(1)", "SS Orig", "SS Bayes", "SS Part 2", "SS Part 4", "SS Part 10",  "Variable") 

CIlenTot %>%
  data.frame() %>%
  gather("Method", "CI.Length", `LME`:`SS.Part.10`) %>%
  mutate(Method = fct_case_when(
    Method == "LME" ~ "LME",
    Method == "LME.AR.1." ~ "LME AR(1)",
    Method == "SS.Orig" ~ "SS Original",
    Method == "SS.Bayes" ~ "SS Bayes",
    Method == "SS.Part.2" ~ "SS Partition k = 2",
    Method == "SS.Part.4" ~ "SS Partition k = 4",
    Method == "SS.Part.10" ~ "SS Partition k = 10"
  )) %>%
  filter(Variable == "B1")%>%
      group_by(Method) %>%
  filter(!is.na(CI.Length)) %>%
  filter(CI.Length < quantile(CI.Length, .99) & CI.Length > quantile(CI.Length, .01)) %>%
  ungroup(Method) %>%
  ggplot(aes(x = Method, y = CI.Length, fill = Method)) + 
  geom_violin() + 
  stat_summary(fun=median, geom="point", shape=18, size=5) +
  facet_grid(Variable ~ .) +
  theme(axis.text.x = element_text(angle = -45, vjust = .99, hjust = 0), legend.position = "none")
```

















```{r, results='asis'}
find <- 6
Itout <- readRDS(paste("Simulations/SimOutReform/",flist[find], sep = "")) 
MedEst <- apply(Itout$Estimate, 1:2, median, na.rm = TRUE)
MedBias <- MedEst - B
VarEst <- apply(Itout$Estimate, 1:2, var, na.rm = TRUE)
Coverage <- apply(Itout$UCL >= B & Itout$LCL <= B, 1:2, mean, na.rm = TRUE) 
colnames(MedBias) <- colnames(VarEst) <- colnames(Coverage) <- c("LME", "LME AR(1)", "SS Orig", "SS Bayes", "SS Part 2", "SS Part 4", "SS Part 10") 

vvec <- flist[find] %>% 
  gsub("Itout ", "", .) %>%
  gsub(".RDS", "", .) %>%
  str_split(" ") %>%
  .[[1]] %>%
  gsub("_", ".", .)


```

## $\sigma^2_\varepsilon =$ `r vvec[1]`, $\sigma^2_\eta =$ `r vvec[2]`

### Coverage

```{r}
Coverage %>%
  kable() %>% kable_styling()
```


### Bias Plot

```{r, fig.height=7}
Bias <- Itout$Estimate - B
Var1Bias <- t(Bias[1,,])
Var2Bias <- t(Bias[2,,])
Var3Bias <- t(Bias[3,,])
VarBias <- data.frame(rbind(Var1Bias, Var2Bias, Var3Bias))

VarBias$Variable <- sort(rep(paste("B", 1:3, sep = ""), dim(Bias)[3]))



colnames(VarBias) <-  c("LME", "LME AR(1)", "SS Orig", "SS Bayes", "SS Part 2", "SS Part 4", "SS Part 10",  "Variable") 

VarBias %>%
  data.frame() %>%
  gather("Method", "Bias", `LME`:`SS.Part.10`) %>%
  mutate(Method = fct_case_when(
    Method == "LME" ~ "LME",
    Method == "LME.AR.1." ~ "LME AR(1)",
    Method == "SS.Orig" ~ "SS Original",
    Method == "SS.Bayes" ~ "SS Bayes",
    Method == "SS.Part.2" ~ "SS Partition k = 2",
    Method == "SS.Part.4" ~ "SS Partition k = 4",
    Method == "SS.Part.10" ~ "SS Partition k = 10"
  ))%>%
  filter(Variable == "B1") %>%
    group_by(Method) %>%
  filter(Bias < quantile(Bias, .99) & Bias > quantile(Bias, .01)) %>%
  ungroup(Method) %>%
  ggplot(aes(x = Method, y = Bias, fill = Method)) + 
  geom_violin() +
  stat_summary(fun=median, geom="point", shape=18, size=5) +
  facet_grid(Variable ~ .) +
  theme(axis.text.x = element_text(angle = -45, vjust = .99, hjust = 0), legend.position = "none")
```

### CI Length Plot

```{r, fig.height=7}
CIlen <- Itout$UCL -Itout$LCL
CIlen1 <- t(CIlen[1,,])
CIlen2 <- t(CIlen[2,,])
CIlen3 <- t(CIlen[3,,])
CIlenTot <- data.frame(rbind(CIlen1, CIlen2, CIlen3))

CIlenTot$Variable <- sort(rep(paste("B", 1:3, sep = ""), dim(Bias)[3]))



colnames(CIlenTot) <-  c("LME", "LME AR(1)", "SS Orig", "SS Bayes", "SS Part 2", "SS Part 4", "SS Part 10",  "Variable") 

CIlenTot %>%
  data.frame() %>%
  gather("Method", "CI.Length", `LME`:`SS.Part.10`) %>%
  mutate(Method = fct_case_when(
    Method == "LME" ~ "LME",
    Method == "LME.AR.1." ~ "LME AR(1)",
    Method == "SS.Orig" ~ "SS Original",
    Method == "SS.Bayes" ~ "SS Bayes",
    Method == "SS.Part.2" ~ "SS Partition k = 2",
    Method == "SS.Part.4" ~ "SS Partition k = 4",
    Method == "SS.Part.10" ~ "SS Partition k = 10"
  )) %>%
  filter(Variable == "B1")%>%
  filter(!is.na(CI.Length)) %>%
      group_by(Method) %>%
  filter(CI.Length < quantile(CI.Length, .99) & CI.Length > quantile(CI.Length, .01)) %>%
  ungroup(Method) %>%
  ggplot(aes(x = Method, y = CI.Length, fill = Method)) + 
  geom_violin() + 
  stat_summary(fun=median, geom="point", shape=18, size=5) +
  facet_grid(Variable ~ .) +
  theme(axis.text.x = element_text(angle = -45, vjust = .99, hjust = 0), legend.position = "none")
```















```{r, results='asis'}
find <-7
Itout <- readRDS(paste("Simulations/SimOutReform/",flist[find], sep = "")) 
MedEst <- apply(Itout$Estimate, 1:2, median, na.rm = TRUE)
MedBias <- MedEst - B
VarEst <- apply(Itout$Estimate, 1:2, var, na.rm = TRUE)
Coverage <- apply(Itout$UCL >= B & Itout$LCL <= B, 1:2, mean, na.rm = TRUE) 
colnames(MedBias) <- colnames(VarEst) <- colnames(Coverage) <- c("LME", "LME AR(1)", "SS Orig", "SS Bayes", "SS Part 2", "SS Part 4", "SS Part 10") 

vvec <- flist[find] %>% 
  gsub("Itout ", "", .) %>%
  gsub(".RDS", "", .) %>%
  str_split(" ") %>%
  .[[1]] %>%
  gsub("_", ".", .)


```

## $\sigma^2_\varepsilon =$ `r vvec[1]`, $\sigma^2_\eta =$ `r vvec[2]`

### Coverage

```{r}
Coverage %>%
  kable() %>% kable_styling()
```


### Bias Plot

```{r, fig.height=7}
Bias <- Itout$Estimate - B
Var1Bias <- t(Bias[1,,])
Var2Bias <- t(Bias[2,,])
Var3Bias <- t(Bias[3,,])
VarBias <- data.frame(rbind(Var1Bias, Var2Bias, Var3Bias))

VarBias$Variable <- sort(rep(paste("B", 1:3, sep = ""), dim(Bias)[3]))



colnames(VarBias) <-  c("LME", "LME AR(1)", "SS Orig", "SS Bayes", "SS Part 2", "SS Part 4", "SS Part 10",  "Variable") 

VarBias %>%
  data.frame() %>%
  gather("Method", "Bias", `LME`:`SS.Part.10`) %>%
  mutate(Method = fct_case_when(
    Method == "LME" ~ "LME",
    Method == "LME.AR.1." ~ "LME AR(1)",
    Method == "SS.Orig" ~ "SS Original",
    Method == "SS.Bayes" ~ "SS Bayes",
    Method == "SS.Part.2" ~ "SS Partition k = 2",
    Method == "SS.Part.4" ~ "SS Partition k = 4",
    Method == "SS.Part.10" ~ "SS Partition k = 10"
  ))%>%
  filter(Variable == "B1") %>%
    group_by(Method) %>%
  filter(Bias < quantile(Bias, .99) & Bias > quantile(Bias, .01)) %>%
  ungroup(Method) %>%
  ggplot(aes(x = Method, y = Bias, fill = Method)) + 
  geom_violin() +
  stat_summary(fun=median, geom="point", shape=18, size=5) +
  facet_grid(Variable ~ .) +
  theme(axis.text.x = element_text(angle = -45, vjust = .99, hjust = 0), legend.position = "none")
```

### CI Length Plot

```{r, fig.height=7}
CIlen <- Itout$UCL -Itout$LCL
CIlen1 <- t(CIlen[1,,])
CIlen2 <- t(CIlen[2,,])
CIlen3 <- t(CIlen[3,,])
CIlenTot <- data.frame(rbind(CIlen1, CIlen2, CIlen3))

CIlenTot$Variable <- sort(rep(paste("B", 1:3, sep = ""), dim(Bias)[3]))



colnames(CIlenTot) <-  c("LME", "LME AR(1)", "SS Orig", "SS Bayes", "SS Part 2", "SS Part 4", "SS Part 10",  "Variable") 

CIlenTot %>%
  data.frame() %>%
  gather("Method", "CI.Length", `LME`:`SS.Part.10`) %>%
  mutate(Method = fct_case_when(
    Method == "LME" ~ "LME",
    Method == "LME.AR.1." ~ "LME AR(1)",
    Method == "SS.Orig" ~ "SS Original",
    Method == "SS.Bayes" ~ "SS Bayes",
    Method == "SS.Part.2" ~ "SS Partition k = 2",
    Method == "SS.Part.4" ~ "SS Partition k = 4",
    Method == "SS.Part.10" ~ "SS Partition k = 10"
  )) %>%
  filter(Variable == "B1")%>%
  filter(!is.na(CI.Length)) %>%
      group_by(Method) %>%
  filter(CI.Length < quantile(CI.Length, .99) & CI.Length > quantile(CI.Length, .01)) %>%
  ungroup(Method) %>%
  ggplot(aes(x = Method, y = CI.Length, fill = Method)) + 
  geom_violin() + 
  stat_summary(fun=median, geom="point", shape=18, size=5) +
  facet_grid(Variable ~ .) +
  theme(axis.text.x = element_text(angle = -45, vjust = .99, hjust = 0), legend.position = "none")
```






```{r, results='asis'}
find <-8
Itout <- readRDS(paste("Simulations/SimOutReform/",flist[find], sep = "")) 
MedEst <- apply(Itout$Estimate, 1:2, median, na.rm = TRUE)
MedBias <- MedEst - B
VarEst <- apply(Itout$Estimate, 1:2, var, na.rm = TRUE)
Coverage <- apply(Itout$UCL >= B & Itout$LCL <= B, 1:2, mean, na.rm = TRUE) 
colnames(MedBias) <- colnames(VarEst) <- colnames(Coverage) <- c("LME", "LME AR(1)", "SS Orig", "SS Bayes", "SS Part 2", "SS Part 4", "SS Part 10") 

vvec <- flist[find] %>% 
  gsub("Itout ", "", .) %>%
  gsub(".RDS", "", .) %>%
  str_split(" ") %>%
  .[[1]] %>%
  gsub("_", ".", .)


```

## $\sigma^2_\varepsilon =$ `r vvec[1]`, $\sigma^2_\eta =$ `r vvec[2]`

### Coverage

```{r}
Coverage %>%
  kable() %>% kable_styling()
```


### Bias Plot

```{r, fig.height=7}
Bias <- Itout$Estimate - B
Var1Bias <- t(Bias[1,,])
Var2Bias <- t(Bias[2,,])
Var3Bias <- t(Bias[3,,])
VarBias <- data.frame(rbind(Var1Bias, Var2Bias, Var3Bias))

VarBias$Variable <- sort(rep(paste("B", 1:3, sep = ""), dim(Bias)[3]))



colnames(VarBias) <-  c("LME", "LME AR(1)", "SS Orig", "SS Bayes", "SS Part 2", "SS Part 4", "SS Part 10",  "Variable") 

VarBias %>%
  data.frame() %>%
  gather("Method", "Bias", `LME`:`SS.Part.10`) %>%
  mutate(Method = fct_case_when(
    Method == "LME" ~ "LME",
    Method == "LME.AR.1." ~ "LME AR(1)",
    Method == "SS.Orig" ~ "SS Original",
    Method == "SS.Bayes" ~ "SS Bayes",
    Method == "SS.Part.2" ~ "SS Partition k = 2",
    Method == "SS.Part.4" ~ "SS Partition k = 4",
    Method == "SS.Part.10" ~ "SS Partition k = 10"
  ))%>%
  filter(Variable == "B1") %>%
    group_by(Method) %>%
  filter(Bias < quantile(Bias, .99) & Bias > quantile(Bias, .01)) %>%
  ungroup(Method) %>%
  ggplot(aes(x = Method, y = Bias, fill = Method)) + 
  geom_violin() +
  stat_summary(fun=median, geom="point", shape=18, size=5) +
  facet_grid(Variable ~ .) +
  theme(axis.text.x = element_text(angle = -45, vjust = .99, hjust = 0), legend.position = "none")
```

### CI Length Plot

```{r, fig.height=7}
CIlen <- Itout$UCL -Itout$LCL
CIlen1 <- t(CIlen[1,,])
CIlen2 <- t(CIlen[2,,])
CIlen3 <- t(CIlen[3,,])
CIlenTot <- data.frame(rbind(CIlen1, CIlen2, CIlen3))

CIlenTot$Variable <- sort(rep(paste("B", 1:3, sep = ""), dim(Bias)[3]))



colnames(CIlenTot) <-  c("LME", "LME AR(1)", "SS Orig", "SS Bayes", "SS Part 2", "SS Part 4", "SS Part 10",  "Variable") 

CIlenTot %>%
  data.frame() %>%
  gather("Method", "CI.Length", `LME`:`SS.Part.10`) %>%
  mutate(Method = fct_case_when(
    Method == "LME" ~ "LME",
    Method == "LME.AR.1." ~ "LME AR(1)",
    Method == "SS.Orig" ~ "SS Original",
    Method == "SS.Bayes" ~ "SS Bayes",
    Method == "SS.Part.2" ~ "SS Partition k = 2",
    Method == "SS.Part.4" ~ "SS Partition k = 4",
    Method == "SS.Part.10" ~ "SS Partition k = 10"
  )) %>%
  filter(Variable == "B1")%>%
  filter(!is.na(CI.Length)) %>%
    group_by(Method) %>%
  filter(CI.Length < quantile(CI.Length, .99) & CI.Length > quantile(CI.Length, .01)) %>%
  ungroup(Method) %>%
  ggplot(aes(x = Method, y = CI.Length, fill = Method)) + 
  geom_violin() + 
  stat_summary(fun=median, geom="point", shape=18, size=5) +
  facet_grid(Variable ~ .) +
  theme(axis.text.x = element_text(angle = -45, vjust = .99, hjust = 0), legend.position = "none")
```