---
output: pdf_document
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
---

```{r, include = FALSE}
library(tidyverse)
library(lme4)
library(nlme)
library(gridExtra)
library(knitr)
library(kableExtra)
library(latex2exp)
library(abind)
```


```{r, cache = TRUE, include = FALSE}

knitr::opts_chunk$set(
    # This should allow Rmarkdown to locate the data
    root.dir = rprojroot::find_rstudio_root_file()
)
# load("NACC/Data/RData")
# rm(list = ls()[ls() != "longdata"])
fct_case_when <- function(...) {
  args <- as.list(match.call())
  levels <- sapply(args[-1], function(f) f[[3]])  # extract RHS of formula
  levels <- levels[!is.na(levels)]
  factor(dplyr::case_when(...), levels=levels)
}
#Bins for histogram
base_bins <- function(type) {
  fun <- switch(type,
    Sturges = nclass.Sturges,
    scott = nclass.scott,
    FD = nclass.FD,
    stop("Unknown type", call. = FALSE)
  )
  
  function(x) {
    (max(x) - min(x)) / fun(x)
  }
}

#Functions to remove missing code in tests
m1 <- function(x){
  ifelse(x %in% c(95, 96, 97, 98, -4), NA, x)
}
m2 <- function(x){
  ifelse(x %in% c(995, 996, 997, 998, -4), NA, x)
}

mmap <- function(data, label, column,...){
  data %>%
    mutate(label = map(column, ...))
}


mmap <- function(data, label, column,...){
  label <- rlang::ensym(label)
  data %>%
    mutate(!!label := map(!!!rlang::enquos(column), ...))
}

mmap2 <- function(data, label, column1, column2,...){
  label <- rlang::ensym(label)
  data %>%
    mutate(!!label := map2(!!!rlang::enquos(column1), !!!rlang::enquos(column2), ...))
}


# Read in NACC data
longdata1 <- read.csv("NACC/Data2/investigator_nacc49.csv")


longdata <- longdata1 %>%
  #Select variables needed for this analysis
  select(
    NACCID, NACCUDSD, NACCFDYS, VISITYR, BIRTHYR, RACE, EDUC, SEX, NACCAPOE, NACCALZD,
    LOGIMEM, MEMUNITS, ANIMALS, VEG, BOSTON, TRAILA, TRAILB, DIGIF, DIGIB, WAIS
  ) %>%
  #Remove missing codes in tests
  mutate_at(vars(LOGIMEM, MEMUNITS, DIGIF, DIGIB, ANIMALS, VEG, WAIS, BOSTON), m1) %>%
  mutate_at(vars(TRAILA, TRAILB), m2) %>%
  #Select only those who are demented for each observation
  group_by(NACCID) %>%
  arrange(NACCID,NACCFDYS)



```






```{r, cache = TRUE}
source("functions/BayesKalmUneq2.R") 

blist <- data.frame(
  # Outcome = c("LOGIMEM", "MEMUNITS", "ANIMALS", "VEG", "BOSTON", "TRAILA", "TRAILB", "DIGIF", "DIGIB", "WAIS")
  Outcome = c("TRAILA", "TRAILB")
) %>%
  mutate(ndata = map(Outcome, function(outcome){
    longdata$Outcome <- longdata[[outcome]]
 
     longdata %>%
      filter(
        !is.na(Outcome), !is.na(VISITYR), !is.na(BIRTHYR), !is.na(RACE), !is.na(EDUC), !is.na(SEX), !is.na(NACCAPOE)
      ) %>%
      arrange(NACCID, NACCFDYS) %>%
      #Select those who started as normal and transitioned to impairment
      filter(NACCUDSD[1] == 1 & any(NACCUDSD %in% c(3,4)), any(NACCALZD == 1)) %>%
      #Derive additional variables
      mutate(
        time = NACCFDYS/365.25, 
        AGE = VISITYR - BIRTHYR,
        AgeBase = AGE[1],
        RACEWHITE = ifelse(RACE == 1, 1, 0),
        APOE = ifelse(NACCAPOE %in% c(2, 4), 1, ifelse(NACCAPOE == 9, NA, 0)),
        SEX = SEX - 1,
        APOESEX = APOE * (SEX),
        Intercept = 1,
        #Do they have an impairment
        MCIoD = NACCUDSD %in% c(3,4),
        FT = min(time[MCIoD]),
        #All time past initial impairment
        DEC = time >= FT,
        #Did the subject ever go to impairment to non-impairment
        reverter = MCIoD != DEC
      ) %>%
      arrange(NACCID, AGE)%>%
      #Remove all reverters
      filter(!any(reverter)) %>%
      #Select all variables that will be used for modeling
      select(NACCID, time,Intercept, SEX, EDUC, RACEWHITE, AgeBase, DEC, APOE, APOESEX, NACCUDSD,
        Outcome) %>%
      na.omit() %>%
      group_by(NACCID) %>%
      filter(any(DEC >0) & any(DEC == 0)) %>%
      #Create knot for those when they transitioned
      mutate(DEC = DEC * (time > 0) * (time - min(time[DEC > 0]))) %>%
      ungroup(NACCID) %>%
      mutate(A1 = AgeBase, E1 = EDUC, AgeBase = AgeBase - mean(AgeBase), EDUC = EDUC - mean(EDUC)) %>%
      # mutate(SEX = SEX * time, RACEWHITE = RACEWHITE * time, APOE = (APOE*2-1) * time, APOESEX = (APOESEX *2 -1), EDUC = EDUC * time, AgeBase = AgeBase * time)
      mutate(SEX = SEX * time, RACEWHITE = RACEWHITE * time, APOE = APOE * time, APOESEX = APOESEX *time, EDUC = EDUC * time, AgeBase = AgeBase * time) 
  }))

```


```{r}

blist <- blist %>%
  mmap(coefs, ndata, function(ld2)coef(lm(Outcome ~ (time + SEX + EDUC + RACEWHITE + AgeBase + DEC + APOE + APOESEX), data = ld2)))

blist <- blist %>%
  mmap2(bayesout, ndata, coefs, function(ld2, Beta.Initial){
    set.seed(1234)
    Its <- 2000
    BayesKalm.Uneq(
      y.long = ld2$Outcome, X.long = as.matrix(ld2[,c( "time", "SEX", "EDUC", "RACEWHITE", "AgeBase", "DEC", "APOE", "APOESEX")]), id = ld2$NACCID, time = ld2$time,
      Burn = Burn, Its = Its, 
      Beta.Initial = Beta.Initial[-1], sigma2.beta = 20, 
      u0 = 0, P0 = 100
    )
  })
# saveRDS(blist, "blist.RDS")
```

```{r}
# blist <- readRDS("blist.RDS")
```


```{r}
blist <- blist %>%
  mmap(bayesbeta, bayesout, function(bkout){

    bdf <- bkout$Beta %>%
      t() %>%
      data.frame() 
    colnames(bdf) <- c( "time", "SEX", "EDUC", "RACEWHITE", "AgeBase", "DEC", "APOE", "APOESEX")
    bdf %>%
      mutate(ind = 1:nrow(.)) %>%
      gather("Beta", "Value", -ind)
  }) 

```


```{r, eval = FALSE}
Burn <- 500
blist <- blist %>%
  mmap(plots, bayesbeta, function(betadf){
    betadf %>%
      filter(ind > Burn) %>%
      group_by(Beta) %>%
      mutate(LCL = quantile(Value, .025), UCL = quantile(Value, .975), Mean = mean(Value)) %>%
      filter(Beta %in% c("APOE", "APOESEX")) %>%
      ggplot(aes(x = ind, y = Value, color = Beta)) + 
      geom_point(shape = 21) + 
      facet_wrap(. ~ Beta, scales = "free_y") + 
      geom_vline(aes(xintercept = Burn)) +
      geom_hline(aes(yintercept = 0), alpha = .65, size = 1.5) +
      geom_hline(aes(yintercept = LCL), color = "red", type = 2) +
      geom_hline(aes(yintercept = UCL), color = "red", type = 2) +
      geom_hline(aes(yintercept = Mean), color = "blue") +
      theme(legend.position = "none")
  }) %>%
  mmap2(plots, plots, Outcome, ~.x + ggtitle(.y))


```



```{r, fig.height=20}
Burn <- 500
blist <- blist %>%
  mmap(bayesestimates, bayesbeta, function(x){
    x %>%
      filter(ind > Burn) %>%
      group_by(Beta) %>%
      summarise(effect = mean(Value), LCL = quantile(Value, .025), UCL = quantile(Value, .975), sig = LCL > 0 | UCL < 0)
  })

blist %>%
  select(Outcome, bayesestimates) %>%
  unnest() %>%
  arrange(Outcome, Beta)%>%
  mutate(CIlength = UCL - LCL)


  
```




```{r}
blist <- blist %>%
  mmap(varplots, bayesout, function(bkout){
    data.frame(bkout$sigma2.eps, bkout$sigma2.eta) %>%
      set_names(c("var.eps", "var.eta")) %>%
      mutate(ind = 1:nrow(.)) %>%
      filter(ind > Burn) %>%
      gather("Variance", "Value", var.eps, var.eta) %>%
      group_by(Variance) %>%
      mutate(LCL = quantile(Value, .025), UCL = quantile(Value, .975), Mean = mean(Value)) %>%
      ggplot(aes(x = ind, y = Value, color = Variance)) + 
      geom_point(shape = 21) + 
      facet_wrap(. ~ Variance, scales = "free_y") + 
      geom_vline(aes(xintercept = Burn)) +
      geom_hline(aes(yintercept = LCL), color = "red", type = 2) +
      geom_hline(aes(yintercept = UCL), color = "red", type = 2) +
      geom_hline(aes(yintercept = Mean), color = "blue") +
      theme(legend.position = "none")
  }) %>%
  mmap2(varplots, varplots, Outcome, ~.x + ggtitle(.y))


```


```{r}
data %>%
  group_by(id) %>%
  filter()
```




```{r}
blist <- blist %>%
  mmap(varresults, bayesout, function(bkout){
    data.frame(bkout$sigma2.eps, bkout$sigma2.eta) %>%
      set_names(c("var.eps", "var.eta")) %>%
      mutate(ind = 1:nrow(.)) %>%
      filter(ind > Burn) %>%
      gather("Variance", "Value", var.eps, var.eta) %>%
      group_by(Variance) %>%
      summarise(LCL = quantile(Value, .025), UCL = quantile(Value, .975), Mean = mean(Value))
  })
blist$varresults
```



```{r}
blist <- blist %>%
  mmap(lmemod, ndata, function(ld2){
    lmeMod <- lme(Outcome ~ (time +SEX + EDUC + RACEWHITE + AgeBase + DEC + APOE + APOESEX), random =  ~1|NACCID, data = ld2, method = "ML", control = lmeControl(opt = "optim"))
  })

blist <- blist %>%
  mmap(lmesum, lmemod, function(mod){
    intervals(mod) %>%
    pluck("fixed") %>%
  as.data.frame() %>%
  rownames_to_column("Beta") %>%
      select(Beta, effect = est., LCL = lower, UCL = upper) %>%
      mutate(sig = LCL > 0 | UCL < 0)
  })


blist %>%
  select(Outcome, lmesum) %>%
  unnest() %>%
  filter(Beta %in% c("APOE", "APOESEX"))
```








```{r, cache=TRUE, include = FALSE}
lmeMod <- lme(ANIMALS ~ (time +SEX + EDUC + RACEWHITE + AgeBase + DEC + APOE + APOESEX), random =  ~1|NACCID, data = ld2, method = "ML", control = lmeControl(opt = "optim"))



lmeDat <- ld2
lmeDat$PredVal <- predict(lmeMod, lmeDat)
```



```{r, cache=TRUE, include = FALSE}
lmeARMod <- lme(ANIMALS ~ (time +SEX + EDUC + RACEWHITE + AgeBase + DEC + APOE + APOESEX), correlation = corAR1(form = ~time), random =  ~1|NACCID, data = ld2, method = "ML")



lmeARDat <- ld2
lmeARDat$PredVal <- predict(lmeARMod, lmeARDat)
```


```{r, cache=TRUE, include = FALSE, eval = TRUE}
source("functions/KalmanRegression2.R")
source("functions/KalmanRecReform2.R")



kp <- KalmanReg2(
  y.long = ld2$ANIMALS, X.long = as.matrix(ld2[,c( "time", "SEX", "EDUC", "RACEWHITE", "AgeBase", "DEC", "APOE", "APOESEX")]), 
  id = ld2$NACCID, time = ld2$time, 
  Beta.Initial = Beta.Initial[-1], a1 = 0, P1 = 1e7, k = 25
)


```


# NACC APOE e4 Data Analysis

Using the described data from the NACC, we fit the model of interest with a random intercept LMEM, an LMEM with a random intercept and an AR(1) temporal variance structure, a partitioned LLT with a group size of 50 in each partition, and a Bayesian LLT. The primary question is whether those with the presence of the APOE e4 allele have a different rate of decline than those without an APE e4 allele. As sex may an interactive effect with APOE e4 on cognition, the interaction variable is also a parameter of interest. We test the null hypothesis of no effect of the APOE e4 allele on cognitive trajectory at the 0.05 significance level.

## NACC Analysis Results

The effect estimates of the linear effect of APOE e4 on the Animals test trajectory are consistent with both simulation studies. The male participant APOE e4 effect estimate confidence intervals are smaller for the random intercept LMEM (0.085) and the AR(1) random intercept LMEM (0.219) than the partitioned LLT (0.360) and the Bayesian LLT (0.280). As the Real Data Simulation suggest, the smaller confidence interval for the LMEMs indicates an improper Type I error. 


```{r effectTab}


ModEs <- rbind(

intervals(lmeMod)$fixed[c(8,9),] %>%
  apply(1, function(x){
    x <- round(x, 3)
    paste(x[2], " (", x[1], ", ", x[3], ")", sep = "")
  }),

intervals(lmeARMod)$fixed[c(8,9),] %>%
  apply(1, function(x){
    x <- round(x, 3)
    paste(x[2], " (", x[1], ", ", x[3], ")", sep = "")
  }),
  kp$summary[7:8,] %>%
  apply(1, function(x){
    x <- round(x, 3)
    paste(x[1], " (", x[2], ", ", x[3], ")", sep = "")
  }),  
rbind(
  apply(bkout$Beta[,Burn:Its], 1, mean)[c(7, 8)],
  apply(bkout$Beta[,Burn:Its], 1, quantile, c(.025, .975))[,c(7, 8)]
) %>%
  apply(2, function(x){
    x <- round(x, 3)
    paste(x[1], " (", x[2], ", ", x[3], ")", sep = "")
  })
)

colnames(ModEs) <- c("APOE", "APOE x Sex")

row.names(ModEs) <- c("LMEM", "LMEM AR(1)", "Partition LLT", "Bayesian LLT")

ModEs%>%
  kbl(longtable = TRUE, escape = FALSE, caption = "Effect estimates for APOE and APOE x Sex interaction") %>%
  kable_styling()


```

```{r,eval = FALSE}
vardf <- data.frame(
  draw = 1001:2000,
  sigeps = bkout$sigma2.eps[1001:2000],
  sigeta = bkout$sigma2.eta[1001:2000]
) %>%
  gather("Param", "Value", sigeps, sigeta) %>%
  mutate(
    Param2 = fct_case_when(
      Param == "sigeps" ~ "$\\sigma^2_{\\epsilon}$",
      Param == "sigeta" ~ "$\\sigma^2_\\eta$"
    )
  )

levels(vardf$Param2) <- TeX(levels(vardf$Param2))

vardf %>%
  ggplot(aes(x = draw, y = Value, color = Param2)) +
  geom_point() +
  facet_wrap(Param2 ~ ., scale = "free_y",
  labeller=label_parsed) +
  stat_summary(aes(x = 1001, yintercept = ..y..), fun=mean, geom="hline", color = "black") +
  theme(legend.position = "none")
```


```{r, eval = FALSE}
VarData <- data.frame(
  draw = 1001:2000,
  sigeps = bkout$sigma2.eps[1001:2000],
  sigeta = bkout$sigma2.eta[1001:2000]
) %>%
  gather("Param", "Value", sigeps, sigeta) %>%
  group_by(Param) %>%
  summarise(
    ParamOut = paste(
      round(mean(Value), 3), 
      " (", 
      paste(round(quantile(Value, c(.025, 0.975)), 3), collapse = ", "), ")", 
      sep = ""
    )
  ) %>%
  mutate(
    Param2 = fct_case_when(
      Param == "sigeps" ~ "$\\sigma^2_{\\epsilon}$",
      Param == "sigeta" ~ "$\\sigma^2_\\eta$"
    )
  )

levels(VarData$Param2) <- TeX(levels(VarData$Param2))

kbl(VarData, row.names = FALSE,  escape = FALSE) %>%
  kable_styling()
```

