---
output: pdf_document
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
---

```{r, include = FALSE}
library(tidyverse)
library(lme4)
library(nlme)
library(gridExtra)
library(knitr)
library(kableExtra)
library(latex2exp)
library(abind)
```


```{r, cache = TRUE, include = FALSE}

knitr::opts_chunk$set(
    # This should allow Rmarkdown to locate the data
    root.dir = rprojroot::find_rstudio_root_file()
)
# load("NACC/Data/RData")
# rm(list = ls()[ls() != "longdata"])
fct_case_when <- function(...) {
  args <- as.list(match.call())
  levels <- sapply(args[-1], function(f) f[[3]])  # extract RHS of formula
  levels <- levels[!is.na(levels)]
  factor(dplyr::case_when(...), levels=levels)
}
#Bins for histogram
base_bins <- function(type) {
  fun <- switch(type,
    Sturges = nclass.Sturges,
    scott = nclass.scott,
    FD = nclass.FD,
    stop("Unknown type", call. = FALSE)
  )
  
  function(x) {
    (max(x) - min(x)) / fun(x)
  }
}

#Functions to remove missing code in tests
m1 <- function(x){
  ifelse(x %in% c(95, 96, 97, 98, -4), NA, x)
}
m2 <- function(x){
  ifelse(x %in% c(995, 996, 997, 998, -4), NA, x)
}


# Read in NACC data
longdata1 <- read.csv("NACC/Data2/investigator_nacc49.csv")

longdata <- longdata1 %>%
  #Select variables needed for this analysis
  select(
    NACCID, NACCUDSD, NACCFDYS, VISITYR, BIRTHYR, RACE, EDUC, SEX, NACCAPOE,
    ANIMALS, NACCALZD
  ) %>%
  #Remove missing codes in tests
  mutate_at(vars(ANIMALS), m1) %>%
  #Select only those who are demented for each observation
  group_by(NACCID) %>%
  arrange(NACCID,NACCFDYS) %>%
  filter(
    !is.na(ANIMALS), !is.na(VISITYR), !is.na(BIRTHYR), !is.na(RACE), !is.na(EDUC), !is.na(SEX), !is.na(NACCAPOE)
  ) %>%
  arrange(NACCID, NACCFDYS) %>%
  #Select those who started as normal and transitioned to impairment
  filter(NACCUDSD[1] == 1 & any(NACCUDSD %in% c(3,4)), any(NACCALZD == 1)) %>%
  #Derive additional variables
  mutate(
    time = NACCFDYS/365.25, 
    AGE = VISITYR - BIRTHYR,
    AgeBase = AGE[1],
    RACEWHITE = ifelse(RACE == 1, 1, 0),
    APOE = ifelse(NACCAPOE %in% c(2, 4), 1, ifelse(NACCAPOE == 9, NA, 0)),
    SEX = SEX - 1,
    APOESEX = APOE * (SEX),
    Intercept = 1,
    #Do they have an impairment
    MCIoD = NACCUDSD %in% c(3,4),
    FT = min(time[MCIoD]),
    #All time past initial impairment
    DEC = time >= FT,
    #Did the subject ever go to impairment to non-impairment
    reverter = MCIoD != DEC
  ) %>%
  arrange(NACCID, AGE)%>%
  #Remove all reverters
  filter(!any(reverter)) %>%
  #Select all variables that will be used for modeling
  select(NACCID, time,Intercept, ANIMALS, SEX, EDUC, RACEWHITE, AgeBase, DEC, APOE, APOESEX, NACCUDSD) %>%
  na.omit() %>%
  #Create knot for those when they transitioned
  mutate(DEC = DEC * (time > 0) * (time - min(time[DEC > 0]))) %>%
  ungroup(NACCID) %>%
  mutate(A1 = AgeBase, E1 = EDUC, AgeBase = AgeBase - mean(AgeBase), EDUC = EDUC - mean(EDUC))


knitr::opts_chunk$set(echo = FALSE)


```






```{r, cache = TRUE}
source("functions/BayesKalmUneq2.R") 
ld2 <- longdata  %>%
  # mutate(SEX = SEX * time, RACEWHITE = RACEWHITE * time, APOE = (APOE*2-1) * time, APOESEX = (APOESEX *2 -1), EDUC = EDUC * time, AgeBase = AgeBase * time)
  mutate(SEX = SEX * time, RACEWHITE = RACEWHITE * time, APOE = APOE * time, APOESEX = APOESEX *time, EDUC = EDUC * time, AgeBase = AgeBase * time) 



ld2$Intercept <- 1L
Beta.Initial <- coef(lm(ANIMALS ~ (time + SEX + EDUC + RACEWHITE + AgeBase + DEC + APOE + APOESEX), data = ld2))

```




```{r, cache = TRUE, include = FALSE}
set.seed(1234)

Its <- 2000
Burn <- floor(Its/2)

bkout <- BayesKalm.Uneq(
    y.long = ld2$ANIMALS, X.long = as.matrix(ld2[,c( "time", "SEX", "EDUC", "RACEWHITE", "AgeBase", "DEC", "APOE", "APOESEX")]), id = ld2$NACCID, time = ld2$time,
    Burn = Burn, Its = Its, 
    Beta.Initial = Beta.Initial[-1], sigma2.beta = 20, 
    u0 = 0, P0 = 100
  )

Beta.b <- bkout$Beta[,Burn:Its]
alpha.b <- bkout$mu[,,Burn:Its]

BetaMean <- apply(Beta.b, 1, mean)
alphaMean <- apply(alpha.b, 1:2, mean)



X.b <- bkout$X
time.b <- bkout$timeMat
id.b <- bkout$id
y.b <- bkout$y

XBetaAll <- map(1:dim(X.b)[3],  function(x) X.b[,,x] %*% BetaMean)


Preds <- map(1:dim(X.b)[3],  function(x) X.b[,,x] %*% Beta.b + alpha.b[,x,])

PredVal <- Preds %>%
  map(function(x){
    apply(x, 1, function(y){
      c(mean(y), quantile(y, c(.025, .975)))
    })
  })

TimeMats <- map(1:length(PredVal), function(x){
  cbind(data.frame(
    id = id.b,
    time = time.b[,x],
    y = y.b[,x],
    PredVal = PredVal[[x]][1,],
    PredLCL = PredVal[[x]][2,],
    PredUCL = PredVal[[x]][3,],
    alpha = alphaMean[,x],
    Xbeta = XBetaAll[[x]]
  ), X.b[,,x])
})


CompleteData <- do.call("rbind", TimeMats)

```


```{r, cache=TRUE, include = FALSE}
lmeMod <- lme(ANIMALS ~ (time +SEX + EDUC + RACEWHITE + AgeBase + DEC + APOE + APOESEX), random =  ~1|NACCID, data = ld2, method = "ML", control = lmeControl(opt = "optim"))



lmeDat <- ld2
lmeDat$PredVal <- predict(lmeMod, lmeDat)
```



```{r, cache=TRUE, include = FALSE}
lmeARMod <- lme(ANIMALS ~ (time +SEX + EDUC + RACEWHITE + AgeBase + DEC + APOE + APOESEX), correlation = corAR1(form = ~time), random =  ~1|NACCID, data = ld2, method = "ML")



lmeARDat <- ld2
lmeARDat$PredVal <- predict(lmeARMod, lmeARDat)
```


```{r, cache=TRUE, include = FALSE, eval = TRUE}
source("functions/KalmanRegression2.R")
source("functions/KalmanRecReform2.R")



kp <- KalmanReg2(
  y.long = ld2$ANIMALS, X.long = as.matrix(ld2[,c( "time", "SEX", "EDUC", "RACEWHITE", "AgeBase", "DEC", "APOE", "APOESEX")]), 
  id = ld2$NACCID, time = ld2$time, 
  Beta.Initial = Beta.Initial[-1], a1 = 0, P1 = 1e7, k = 25
)


```



# NACC Analysis Results

The effect estimates of the linear effect of APOE e4 on the Animals test trajectory are consistent with both simulation studies. The male participant APOE e4 effect estimate confidence intervals are smaller for the random intercept LMEM (0.085) and the AR(1) random intercept LMEM (0.219) than the partitioned LLT (0.360) and the Bayesian LLT (0.280). As the Real Data Simulation suggest, the smaller confidence interval for the LMEMs indicates an improper Type I error. 


```{r effectTab}


ModEs <- rbind(

intervals(lmeMod)$fixed[c(8,9),] %>%
  apply(1, function(x){
    x <- round(x, 3)
    paste(x[2], " (", x[1], ", ", x[3], ")", sep = "")
  }),

intervals(lmeARMod)$fixed[c(8,9),] %>%
  apply(1, function(x){
    x <- round(x, 3)
    paste(x[2], " (", x[1], ", ", x[3], ")", sep = "")
  }),
  kp$summary[7:8,] %>%
  apply(1, function(x){
    x <- round(x, 3)
    paste(x[1], " (", x[2], ", ", x[3], ")", sep = "")
  }),  
rbind(
  apply(bkout$Beta[,Burn:Its], 1, mean)[c(7, 8)],
  apply(bkout$Beta[,Burn:Its], 1, quantile, c(.025, .975))[,c(7, 8)]
) %>%
  apply(2, function(x){
    x <- round(x, 3)
    paste(x[1], " (", x[2], ", ", x[3], ")", sep = "")
  })
)

colnames(ModEs) <- c("APOE", "APOE x Sex")

row.names(ModEs) <- c("LMEM", "LMEM AR(1)", "Partition LLT", "Bayesian LLT")

ModEs%>%
  kbl(longtable = TRUE, escape = FALSE, caption = "Effect estimates for APOE and APOE x Sex interaction") %>%
  kable_styling()


```

```{r,eval = FALSE}
vardf <- data.frame(
  draw = 1001:2000,
  sigeps = bkout$sigma2.eps[1001:2000],
  sigeta = bkout$sigma2.eta[1001:2000]
) %>%
  gather("Param", "Value", sigeps, sigeta) %>%
  mutate(
    Param2 = fct_case_when(
      Param == "sigeps" ~ "$\\sigma^2_{\\epsilon}$",
      Param == "sigeta" ~ "$\\sigma^2_\\eta$"
    )
  )

levels(vardf$Param2) <- TeX(levels(vardf$Param2))

vardf %>%
  ggplot(aes(x = draw, y = Value, color = Param2)) +
  geom_point() +
  facet_wrap(Param2 ~ ., scale = "free_y",
  labeller=label_parsed) +
  stat_summary(aes(x = 1001, yintercept = ..y..), fun=mean, geom="hline", color = "black") +
  theme(legend.position = "none")
```


```{r, eval = FALSE}
VarData <- data.frame(
  draw = 1001:2000,
  sigeps = bkout$sigma2.eps[1001:2000],
  sigeta = bkout$sigma2.eta[1001:2000]
) %>%
  gather("Param", "Value", sigeps, sigeta) %>%
  group_by(Param) %>%
  summarise(
    ParamOut = paste(
      round(mean(Value), 3), 
      " (", 
      paste(round(quantile(Value, c(.025, 0.975)), 3), collapse = ", "), ")", 
      sep = ""
    )
  ) %>%
  mutate(
    Param2 = fct_case_when(
      Param == "sigeps" ~ "$\\sigma^2_{\\epsilon}$",
      Param == "sigeta" ~ "$\\sigma^2_\\eta$"
    )
  )

levels(VarData$Param2) <- TeX(levels(VarData$Param2))

kbl(VarData, row.names = FALSE,  escape = FALSE) %>%
  kable_styling()
```

